---
title: "FE 542- HW1"
author: "M.Furkan Isik"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Problem 1

An urn contains three type A coins and two type B coin. When a type A coin is ipped, it comes up heads with probability 1/3, whereas a type B coin is  ipped, it comes up heads with  robability 2/3. A coin is randomly chosen from the urn and ipped. Given that theip landed on heads, what is the probability that it was a type B coin?

Solution:

0.5714286


Prob(B|heads)= P(heads|B).P(B)/(P(heads))

P(heads|B)=if the coin is B prob of getting head is 2/3 which is given in the question

P(heads|A)=if the coin is A prob of getting head is  1/3 which is given in the question

P(B)=2/5 There are totally 5 coins and 2 of them are B coins so the probability of chosing B coin is 2/5

P(A)=3/5 There are totally 5 coins and 3 of them are A coins so the probability of chosing A coin is 3/5


P(heads)= P(heads|B).P(B)+P(heads|A).P(A)



Prob(B|heads)= P(heads|B).P(B)/(P(heads))

- Calculating denominator of the equation
```{r}
denominator=2/3*2/5 + 1/3*3/5
denominator

```


- Calculating numerator of the equation
```{r}
numerator=2/3*2/5
numerator
```

- Showing the result
```{r}
result=numerator/denominator

result


```



### Problem 2

(i) Generate 100 observations from a normal distribution with mean 2 and variance 7.

```{r}
sample=rnorm(100,mean=2,sd=sqrt(7))

sample
```


(ii) Compute the sample mean, standard deviation, skewness and kurtosis (if excess kurtosis state that clearly).


```{r}
mean(sample)
```


```{r}
sd(sample)
```

- Measure of assymetry
- if skewness is positive, the data is skewed right
- if skewness is negative, the data is skewed left
- if skewness is between −1 and −½ or between +½ and +1, the distribution is moderately skewed.
- If skewness is between −½ and +½, the distribution is approximately symmetric.
- If skewness is less than −1 or greater than +1, the distribution is highly skewed.

We can say that sample is approximately symmetric since the skewness value is between −½ and +½
```{r}
library(fBasics)
skewness(sample)
```

- Measure of tail thickness
- Negative excess kurtosis has short tails Platykurtic
- Positive excess kurtosis  have heavy tails Leptokurtic 

We can say that sample is Platykurtic(tails are light)  since excess kurtosis value is negative
```{r}
kurtosis(sample)
```

(iii) Generate a Q-Q plot of the observations from (i) versus the standard normal distribu-tion. Describe your plot in no more than 3 sentences.

-  Data is fairly normal distributed since majority of the data points in line with the 45 degree normal line.
```{r}
qqnorm(sample)
qqline(sample,col="green")
```



### Problem 3

(i) Download daily price data for January 1, 2017 through December 31, 2020 of Microsoft stock from Yahoo Finance. You may use the quantmod package in R for this purpose.


```{r}
library(quantmod)

getSymbols('MSFT',src='yahoo',from='2017-01-01',to='2020-12-31')
```

(ii) Generate a plot of the closing prices and of daily log returns.

- plot of the daily return
```{r}
r_arith=dailyReturn(MSFT)

plot(r_arith)



```


- plot of daily log return
```{r}
r_log = dailyReturn(MSFT,type='log')

plot(r_log)
```

(iii) Compute the sample mean, standard deviation, skewness and kurtosis (if excess kurtosis
state that clearly) of the daily log returns.

```{r}
mean(r_log)
sd(r_log)
skewness(r_log)
kurtosis(r_log)

```

(iv) Perform the Jarque-Bera Normality test (with 5% significance level) for daily log returns
and interpret your results. Compare your result a visualization from a Q-Q plot.

Ho: (Normal Distribution)
Ha: (Non-Normal Distribution)

- p value of Jarque - Bera Normalality Test Normalality Test is very small and smaller than significance level(0.05) so we can reject null hypothesis. Hence, the data is not normally distributed

```{r}
normalTest(r_log,method="jb")

```

(v) Compute and plot the lag-1 through lag-10 autocorrelation of the daily log returns.

- plot the autocorrelation
- calculate the autocorrelation

```{r}
acf(r_log, lag.max=10, plot=TRUE)

acf(r_log, lag.max=10,plot="False")
```


(vi) Test if the lag-5 autocorrelation is non-zero (with 5% significance level). Interpret your
results.

H0 : ρl = 0
Ha : ρl not equal 0
if |t ratio|>Zα/2, we reject null hypothesis



- According to individual ACF test, our t ratio is not bigger than qnorm value. Hence, we fail to reject null.
```{r}

ac = acf(r_log,plot=TRUE)
T = length(r_log)
ratio = sqrt(T)*ac$acf[6]
```


```{r}
significance = 0.05
abs(ratio) > qnorm(1-significance/2) #FALSE = FAIL TO REJECT NULL
```







(vii) Perform the Ljung-Box test to test if the first 2, first 5, and first 8 lagged autocorrela-
tions are non-zero (with 5% significance level). Interpret your results.

H0: The data are independently distributed(no auto correlation)
Ha: The data are not independently distributed; they exhibit serial correlation.


-First two autocorrelations are non zero because  p- value of the Ljung-Box Tests is too small with given 5% significance level. Hence, we can reject null hypothesis(no auto correlation)
```{r}
Box.test(r_log,lag=2,type="Ljung-Box") 
```

- First 5 autocorrelations are non zero because  p- value of the Ljung-Box Tests is too small in comparison to  given 5% significance level. Hence, we can reject null hypothesis( no auto correlation)

```{r}
Box.test(r_log,lag=5,type="Ljung-Box") 
```

- First 8 autocorrelations are non zero because  p- value of the Ljung-Box Tests is too small in comparison to  given 5% significance level. Hence, we can reject null hypothesis( no auto correlation)


```{r}
Box.test(r_log,lag=8,type="Ljung-Box") 
```




