---
title: "FE542-HW5"
author: "M.Furkan Isik"
date: "4/28/2021"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Problem 1


## (i)  Construct single time series autoregressive models for the single time series c1t and c10t.

- To decide order of the model, pacf values with different lags observed
- For c1t we have lowest pacf value at lag 3 and lag 7, therefore lag 2 was chosen for order of the  model
- For c10t we have lowest pacf value at lag 4 and lag 7, therefore lag 3 was chosen for order model


```{r}
df <- read.csv("homework05.csv")
c1t=df$DGS1[2:708]-df$DGS1[1:707]
c10t= df$DGS10[2:708]-df$DGS10[1:707]

ct <- data.frame(c1t,c10t)
c1t.pacf <- pacf(x = c1t, lag.max = 10)
c10t.pacf <- pacf(c10t, lag.max = 10)


```



- Constructing ar models with order 2 and order 3

```{r}
ar1 <- arima(c1t, order = c(2,0,0))
ar10 <- arima(c10t, order = c(3,0,0))
ar1
ar10
```

## ii)  Build a bivariateautoregressive model for the two change series.


- Used Varselect function to find the lowest AIC value
- Lag 2 has the huge drop in AIC value, after that although values are smaller, there is not a huge drop, so lag 2 is chosen

```{r}
library(vars)

ct.var <- VARselect(ct, lag.max = 5)
ct.var$criteria 
```



```{r}
VAR_est <- VAR(ct,2)
summary(VAR_est)
```


-  The reduced form of VAR(1) model is similar to this r(t)=phi(0)+phi(1)*r(t-1)+phi(2)*r(t-1)+a(t)

```{r}

phi0 = c(VAR_est$varresult$c1t$coefficients[5],VAR_est$varresult$c10t$coefficients[5])
phi1 = matrix(c(VAR_est$varresult$c1t$coefficients[1:2],
VAR_est$varresult$c10t$coefficients[1:2]), nrow = 2, byrow = T)
phi2 = matrix(c(VAR_est$varresult$c1t$coefficients[3:4],
VAR_est$varresult$c10t$coefficients[3:4]),nrow = 2, byrow = T)
phi0
phi1
phi2

```


## iii)  Transform the constructed bivariate model into a structural form.

- The bi-variate model will be transfered into structural form
- In order to do that, covariance matrix of  residuals created  and cholesky decompostion applied 
- The VAR(2) structural form would be,  Ar(t)=phi0.star + phi1.star*r(t-1) + phi2.star*r(t-2) + b(t)

```{r}

amat = diag(2)
amat[2,1] = NA
svar1 = SVAR(VAR_est, Amat = amat)
A = svar1$A
B = svar1$B
phi0.star = A%*%phi0
phi1.star = A%*%phi1
phi2.star = A%*%phi2
phi0.star
phi1.star
phi2.star



```


- phi1 value is -0.006801116 
- phi 2 value is 0.02627412 and both phi1 and phi2 is very close to 0.
- it could be said that lag 1 and lag 2 don't have crucial impact on c10t and c1t prediction

```{r}
phi1
phi2

```


## (iv)  Briefly discuss the implications of the vector autoregressive model and compare with the single time series models.

- Both models were compared in terms of Rsquared values 
- R squared value of AR model c1t=0.1846312 and c10t=0.1359105
- R squared value of VAR model  c1t= 0.1978   c10t=0.1321
- Therefore, it could be said that  VAR model is slighly better in comparison to sigle time series model  in terms of  explaning the variability 

```{r}
SST1 = c1t-mean(c1t)
SST1 = SST1^2
SST1 = sum(SST1)
RST1 = (ar1$residuals)^2
RST1 = sum(RST1)
Rsqrd1 <- 1 - (RST1/SST1)
Rsqrd1
```


```{r}
SST2 = c10t-mean(c10t)
SST2 = SST2^2
SST2 = sum(SST2)
RST2 = (ar10$residuals)^2
RST2 = sum(RST2)
Rsqrd2 <- 1 - (RST2/SST2)
Rsqrd2
```




