---
title: "FE 542- HW3"
author: "M.Furkan Isik"
date: "3/22/2021"
output: word_document
---



#Problem 1

In R create a report in pdf format using RMarkdown (or, if you choose to use Python instead,create a Jupyter notebook) to


(i)  Download daily price data for January 1, 1980 through December 31, 2019 of Boeingstock from Yahoo Finance.  You may use the quantmod package in R for this purpose.


```{r}

library(quantmod)


getSymbols(Symbols = "BA",src="yahoo",from="1980-01-01", to="2019-12-31",periodicity="monthly")



```


- Calculating Monthly returns

```{r}
R= BA$BA.Close/lag(BA$BA.Close) -1

R= R[2:length(R)]

head(R)

```


- Calculating Monthly log returns

```{r}

logR=log(1+R)

head(logR)

```


(ii) Is there any evidence of serial correlations in the monthly log returns. Use autocor-
relations and 5% significance level to answer the question. If yes, remove the serial
correlations.


- The p-value of Box-Ljung test is higher than 0.05( 0.3433), so we fail to reject null hypothesis(no serial correlation)

- We could say that there is no serial correlation

```{r}

acf(logR)


Box.test(logR,type="Ljung-Box")

```








(iii) Is there any evidence of ARCH effects in the monthly log returns? Use the residual
series if there are serial correlations in part (ii). Use Ljung-Box statistics for the squared
returns (or residuals) with 6 and 12 lags of autocorrelations and 5% significance level
to answer the question.



- First, find the difference between logreturns and its mean value to come up with residuals

- Then take the square of residuals

```{r}

y= logR-mean(logR)

head(y)


```


- Use Ljung Box test(lag value 6) to test ARCH effects

- The p value of the Ljung-Box statistic is 0.02026, with 5% significance level,  we can say there is ARCH effect in the log returns

```{r}
Box.test(y^2,lag=6,type="Ljung")

```


- Use Ljung Box test(lag value 12) to test ARCH effects


```{r}
Box.test(y^2,lag=12,type="Ljung")
```


(iv) Identify an ARCH model for the data and fit the identified model. Write down the
fitted model and justify your choice of parameters.


- According to PACF and ACF, ARCH(3) model could be appropriate to fit

- Different models tried and modified depending on the significance of coefficients and

- Among ARCH(3,0), ARCH(1,0) and  ARMA(1,0)+ARCH(1,1), ARCH(3,0) model fits the better into the data 


- fitted model Coefficient(s):

       mu      omega     alpha1     alpha2     alpha3  
0.0122236  0.0035564  0.0754545  0.0606447  0.3513973



```{r}

par(mfrow=c(1,2))

pacf(abs(logR^2))

acf(abs(logR^2))


```


- ARCH(3,0) model

```{r}
library(fGarch)

g1=garchFit(logR~garch(3,0),data=logR,trace=F)

summary(g1)

```



- ARCH(1,0) model
- alpha 1 is not significant

```{r}
m1=garchFit(logR~garch(1,0),data=logR,trace=F)

summary(m1)


```



- ARMA(1,0)+ARCH(1,1)

```{r}

# Next, fit an ARMA(1,0)+GARCH(1,1) model with  Gaussian noises

m2=garchFit(logR~arma(1,0)+garch(1,1),data=logR,trace=F)

summary(m2)


```


